---
title: "Transform Covid-19 format"
author: "Chanyanart KiattipornOpas"
date: "10/15/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download Library 
```{r message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
```


## Import Data to R studio 

```{r Import csv file, message=FALSE, warning=FALSE}
confirm <- read_csv("time_series_covid19_confirmed_global.csv")

deaths <- read_csv("time_series_covid19_deaths_global.csv")

recovered <- read_csv("time_series_covid19_recovered_global.csv")

```

## Check Observation / Rows 

```{r check Observation1, echo=TRUE}
confirm <- as_tibble(confirm)
confirm
# 279 Rows*635 Col
```


```{r check Observation2, echo=TRUE}
deaths <- as_tibble(deaths)
deaths
# 279 Rows*635 Col
```


```{r check Observation3, echo=TRUE}
recovered <- as_tibble(recovered)
recovered
# 264 Rows*635 Col 
```
These data are kept with Wide Format, So we need the Long Format for analyzing. 

## Create New Column & Identify File Type 
```{r}
confirm <- confirm %>% mutate(fromfile = "Confirmed") 

deaths <- deaths %>% mutate(fromfile = "Deaths")

recovered <- recovered %>% mutate(fromfile = "Recovered")
```

## Combine Data
```{r}

all_data <- list(confirm, deaths, recovered)
all_data <- bind_rows(all_data)

# 822 rows * 636 Columns
```

## Drop Irrelevent Column 
```{r}
#colnames(all_data)

all_data <- all_data %>% select(-"Province/State",-"Lat", -"Long")

all_data
```

## Group by and Sum 
Data showed duplicated countries where keep data in each area separately (ex. Australia: Victoria, New South Wales), So these data should be combined into 1 number per country per day. 

```{r}
# Rename Column
colnames(all_data)[1] <- "Country"

# Combine data with Group_by()
all_data <- all_data %>% 
            select(Country,
                   fromfile,everything()) %>% 
            group_by(Country,fromfile) %>% 
            summarise_all(sum)

all_data
```

## Transform Wide to Long Format 
```{r}
long <- gather(all_data,-c(Country,fromfile),
                          key = Date,
                          value = Case)

long

# Rename Column 
colnames(long)[2] <- "Type"
```


## Export csv.file 
```{r}
write_csv(long,"/Users/Chanyanart/Desktop/Data/Archived/COVID_19/covid_lastest_date.csv")
```

